---
title: "Data transformation with dplyr"
date: '2021-11-07'
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

##### I import the [data from the website](  https://biostat2.uni.lu/practicals/data/judgments.tsv). 
```{r}
judge_url <- "https://biostat2.uni.lu/practicals/data/judgments.tsv"
judgments <- readr::read_tsv(judge_url)
```
Assign to the name `judgments`


##### I use `glimpse()` to identify columns and column types.
```{r}
glimpse(judgments)
```

##### I select all columns that refer to the STAI questionnaire
```{r}
judgments %>%
  select(starts_with('STAI'))
```

##### I select all subjects older than 25
```{r}
judgments %>%
  filter(age > 25) %>%
  select (subject)
```

##### I retrieve all subjects younger than 20 which are in the stress group
The column for the group is `condition`.
```{r}
judgments %>%
  filter(age < 20, condition == "stress") %>%
  select (subject)
```

##### I abbreviate the gender column such that only the first character remains
```{r}
judgments %>%
  mutate(gender = case_when(
    gender == 'female' ~ 'f',
    gender == 'male' ~ 'm'
  )) %>%
  select(subject, age, gender)
```

##### I normalize the values in the REI group
Divide all entries in the REI questionnaire by 5, the maximal value.
```{r}
judgments %>%
  mutate(across(contains('REI'), `/`, 5)) %>%
  select(contains('REI'))
```

##### I ompute the average and the standard deviation for all moral dilemma columns. 
The resulting column names contain the names of "avg" (for the average) and "sd"
for the standard deviation. 
```{r}
judgments %>%
  summarise(across(starts_with('moral_dil'), 
                   list(avg = ~ mean(.x, na.rm = TRUE),
                        sd = ~ sd(.x, na.rm = TRUE))))
```

