---
author: "Michael Mbajwa"
date: "2021-11-19"
output: html_document
---
  
```{r setup, include = FALSE}
library(tidyverse)
library(broom)
library(purrr)
library(cowplot)
```
  
To practice how to compare samples of a dataset, I am going to use the wings dataset. This dataset contains the measurement of wing length in insects living in 2 different areas.

> Objective: Compare both locations and determine if the insects have different sizes of wings. By _difference_, we mean testing several components of what defines a distribution.



```{r}
# Load the `wings` dataset that contains the insect wing length measured in two different locations.
wings <- read_delim("wings.csv")
wings
```


```{r}

# Brief summary of our dataset
# The variables under the `size` column are continuous - because they take real numbered values. The variables under the `location` column on the other hand are categorical.
```



```{r}
# Before doing anything, we would like to have an idea of how much data we are dealing with: globally and per location. How many observations do we have per location?

wings %>%
  group_by(location) %>%
  summarise(n_obs = n())
```



```{r}
# Now that we have an idea of how much data we are dealing with, let's represent the densities of our datasets graphically.
wings %>%
  filter(location == "A") %>%
  ggplot(aes(x = size)) +
  geom_density(fill = "#e76f51", alpha = 0.8) +
  theme_classic() +
  labs(title = "Density A",
       x = "Size",
       y = "Density") +
  geom_vline(aes(xintercept = mean(size), colour = "mean"), linetype = "dashed", size = 1) +
  theme(legend.title = element_blank()) +
  scale_color_manual(values = c(mean = "black")) -> plot_a
  
wings %>%
  filter(location == "B") %>%
  ggplot(aes(x = size)) +
  geom_density(fill = "#457b9d", alpha = 0.8) +
  theme_classic() +
  labs(title = "Density B",
       x = "Size",
       y = "Density") +
  geom_vline(aes(xintercept = mean(size)), linetype = "dashed", size = 1) -> plot_b

plot_grid(plot_a, plot_b, nrow = 2)
```



```{r}
# To study the medians of the different locations, we can also use a box plot

wings %>%
  filter(location == "A") %>%
  ggplot(aes(x = location,
             y = size)) +
  geom_boxplot(fill = "#e76f51") +
  labs(title = "Box plot for Location A") +
  theme_light()
```


```{r}
wings %>%
  filter(location == "B") %>%
  ggplot(aes(x = location,
             y = size)) +
  geom_boxplot(fill = "#457b9d") +
  labs(title = "Box plot for location B") +
  theme_light()
```


```{r}
# Combining the two boxplots
bp1 <- wings %>%
  filter(location == "A") %>%
  ggplot(aes(x = location,
             y = size)) +
  geom_boxplot(fill = "#e76f51") +
  labs(title = "Box plot for Location A") +
  theme_light()

bp2 <- wings %>%
  filter(location == "B") %>%
  ggplot(aes(x = location,
             y = size)) +
  geom_boxplot(fill = "#457b9d") +
  labs(title = "Box plot for location B") +
  theme_light()

plot_grid(bp1, bp2, nrow = 2)
```



```{r}
# Observing the box plots, we can see the medians of the two locations appear different.
```


## NEXT GOAL: Compare the means of the locations.

Before comparing the two means, we need to ensure that both distributions are normal in order to determine which test we should use. 


```{r}
# To test normality, I will use a shapiro test.
wings %>%
  nest_by(location) %>%
  mutate(norm_test = map(data, shapiro.test),
         tidied = tidy(norm_test)) %>%
  unnest(tidied)
```


```{r}
# Observation: From the output of the result, the p-values are > 0.05 implying that the distributions of locations A and B are not significantly different from the normal distribution. We can assume there is strong evidence A and B follow the normal distribution.
```

Since the samples follow a normal distribution, we will use the t.test() to compare two means.



Before I perform the t.test, I will test if the variances of the sample are the same because it affects my t.test

```{r}

wings %>%
  nest_by(location) %>%
  pivot_wider(names_from = location,
              values_from = data) %>%
  mutate(test = tidy(var.test(unlist(A), unlist(B)))) %>%
  unnest(test)

# Based on the p-value, we can conclude the variances are different between the 2 locations.

# The consequence for the t-test chosen is that we will conduct the test with a welch correction/adjustment.
```
```{r}
# A simple manual test
wings %>%
  group_by(location) %>%
  summarise(variance = var(size),
            mean = mean(size))
```



I will efine $H_0$ for the appropriate test, and then the alternative hypothesis $H_1$.

```{r}
# H_1 = The means of the two locations are equal.
# H_1 = The means of the two locations are unequal.
```


I will then perform the t.test
```{r}
wings %>%
  nest_by(location) %>%
  pivot_wider(names_from = location,
              values_from = data) -> wings_long

t.test(Pair(unlist(A), unlist(B)) ~ 1, data = wings_long, var.equal = FALSE)
```

```{r}
# CONCLUSION
# Judging from the result of our p-value, we can conclude there is a weak evidence that the means are unequal. This mean the means are not significantly different from each other.
```

