---
title: "String manipulation"
author: "Michael Mbajwa"
date: '2021-10-25'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

> This little tutorial aims to make me familiar with some of the functions of the `stringr` package and a few regular expressions. 

## Strings and escape sequences in R

##### Write a sentence with escape sequences.
I try the sentence: `"It's the end of the world!" he said.\` . 
Assign the string to a variable and try as`print()`, `cat()` and `writeLines()`.

```{r}
str_var <- '"It\'s the end of the world!" he said.\\'
print(str_var)
cat('\n')
cat(str_var)
cat('\n')
writeLines(str_var)
```

## `stringr` functions 

We will be using the `words` data that is built into `stringr`. 

```{r}
words[1:50]
length(words)
str_length(words)[1:20]
```


##### Select words that

1. Contain a `y`
```{r}
str_subset(words, 'y')
```

2. Start with `y`
```{r}
str_subset(words, '^y')
```


3. Contain a `y` within the word

```{r}
str_subset(words, '.y.')
```

##### Extract the `y` and the previous character.

Note: Use the function `unique()` around the results to avoid printing many empty matches.

```{r}
unique(str_match(words, '(.{1}y)'))
```


##### Get the lengths of the first ten words

I use `head(words, 10)` as a convenient way to access the elements of the `words` vector.
```{r}
first_ten <- head(words, 10)
str_length(first_ten)
```


## Viral research 

Read the genome sequence of the Hepatitis D virus: [hepd.fasta](https://biostat2.uni.lu/practicals/data/hepd.fasta).

```{r}
hepd_genome <- readr::read_lines("https://biostat2.uni.lu/practicals/data/hepd.fasta")
```

##### I find the length of the genome sequence? 

```{r}
str_length(hepd_genome)
# Length is 1682
```

##### I find the sequence composition and how often each character occur?

```{r}
seq_comp <- unique(str_split(hepd_genome, '')[[1]])
seq_comp
# [1] "A" "T" "G" "C"
str_count(hepd_genome, seq_comp)
# [1] 339 354 485 504
```


##### Find motifs in the sequence using `str_locate()`. 

I find all matches of the sequence $ATG$ in the genome sequence. 

```{r}
str_locate_all(hepd_genome, 'ATG')
```