---
author: "Michael Mbajwa"
date: "2021-11-19"
output: html_document
---


> Perform some explanatory analysis on the penguin dataset provided by the palmer penguins package 


```{r}
# Laod required packages
library(palmerpenguins)
library(tidyverse)
library(ggplot2)
```

```{r}
# Let's have a quick glimpse of the dataset
glimpse(penguins)
```


# I would like to plot the body mass on the _y_ axis and the bill length on the _x_ axis.

```{r}
penguins %>%
  ggplot(aes(x=bill_length_mm,
             y=body_mass_g)) +
  geom_point(aplha=0.7, color = "yellow") +
  labs(x = 'Bill length (mm)',
       y = 'Body mass (g)') +
  theme_dark()
```


# I would plot again the body mass on the _y_ axis and the bill length on the _x_ axis, but different different species with the aid of color.

```{r}
penguins %>%
  ggplot(aes(x=bill_length_mm,
             y=body_mass_g,
             color=species)) +
  geom_point(aplha=0.7) +
  labs(x='Bill length (mm)',
       y='Bodby mass (g)') +
  theme_classic() -> penguin_plot

penguin_plot
```


# Let's add a trend line to the above chart

```{r}
penguin_plot +
  geom_smooth(method = 'lm', formula = 'y ~ x', se = FALSE)
```



# Let's repeat the above plot but this time adding more levels of details (precisely island)

```{r}
# Write your answer here
penguins %>%
  ggplot(aes(x = bill_length_mm,
                 y = body_mass_g,
                 colour = species,
                shape = island)) +
  geom_point(size=3, alpha=0.6) +
  geom_smooth(method = 'lm', formula = 'y ~ x', se = FALSE) +
  scale_color_viridis_d() +
  theme_bw() +
  labs(x='Bill length (mm)',
       y='Bodby mass (g)') +
  theme_classic()
```


# Let's produce a final grand plot


```{r}
# Write your answer here
penguins %>%
  ggplot(aes(x = bill_length_mm,
                 y = body_mass_g,
                 colour = species,
                shape = island)) +
  geom_point(size=3, alpha=0.7) +
  geom_smooth(method = 'lm', formula = 'y ~ x', se = FALSE) +
  scale_color_viridis_d() +
  theme_bw() +
  theme(plot.title = element_text(family="sans"),
        plot.caption = element_text(face = "italic", hjust = 1.5)) +
  labs(x = 'Bill length (mm)',
       y = 'Body mass (g)',
       title = "Penguin bill length and body mass",
       subtitle = 'Dimensions for male/female Adelie, Chinstrap and Gentoo Penguins
       at Palmer Station LTER',
       shape = 'island',
       color = 'Penguin species',
       caption = 'Horst AM, Hill AP, Gorman KB(2020)')
```

