---
author: "Michael Mbajwa"
date: '2021-11-19'
output: pdf_document
---
  
```{r setup, include = FALSE}
library(tidyverse)
library(broom)
library(purrr)
```
  

# The objective of this project is to test for normality of given dataset

I am going to use the `garden` data, with the measurement of ozone concentrations in 3 different areas.

> Objective: the idea is to display the distribution of all three series and look at the shapes. 

## 1.1 Loading the data

```{r}
garden <- read_delim("garden.csv")
garden
```

```{r}
# A little of exploration of plotting a normal distribution
x <- seq(from = -5, to = 5, length.out = 100)
dens_x <- dnorm(x)

ggplot() +
  geom_line(aes(x = x,
                y = dens_x)) +
  labs(title = "A density distribution",
       y = "Density",
       x = " X") + theme_classic()
```


It's now time to apply a density function to the provided data. 

I will start by drawing the density plot of gardenA using `ggplot2`.

```{r}
# Write your answer here
dens_a <- tidy(density(garden$gardenA))

garden %>%
  pivot_longer(everything(),
               names_to = "Garden",
               values_to = "ozone") -> garden_2
garden_2 %>%
  filter(Garden == "gardenA") %>%
  ggplot() +
  geom_density(aes(x = ozone, fill = Garden)) +
  scale_x_continuous(limits = c(min(dens_a$x), max(dens_a$x))) +
  theme_classic() +
  labs(title = "Density plot of Garden A",
       x = "[ozone]",
       y = "density")
```


  
```{r}
# I would like to get limits i.e minimum and maximum of the density from each Garden
garden_2 %>%
  group_by(Garden) %>%
  summarise(min = min(density(ozone)$x),
            max = max(density(ozone)$x)) -> limits_x
limits_x
```


```{r}
# Use the limits from the density, I plot the densities of each garden
garden_2 %>%
  ggplot(aes(x = ozone, color = Garden, fill = Garden)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(min(limits_x$min), max(limits_x$max))) + 
  theme_classic() +
  labs(title = "Garden Densities",
       x = "[ozone]",
       y = "density")
```


```{r}
# Some observations from the density plot

# The empirical distributions of gardenA and gardenB have a bell-shaped curve which
# suggests it follows a normal distribution.
# That of gardenC is a double bell curve. This suggests it's a bimodal distribution.
```


## Plotting Quantile-Quantile comparison to the normal distribution

A Q-Q plot allows you to see how the quantiles of two distributions fit. In our case, we want to compare each of the garden’s distribution to the normal distribution. 

METHOD: I use the function `qqnorm()` and add the ideal line with `qqline()`. When the dots are close to the plain line, then the distribution follows the normal one.

```{r}
# Write your answer here
# Q-Q plot for garden A
qqnorm(garden$gardenA, main = "Normal Q-Q Plot for GardenA")
qqline(garden$gardenA)

# Observation
# The dots are close to the plain line. Means the distribution follows the normal one.
```
```{r}
# Q-Q plot for garden B
qqnorm(garden$gardenB, main = "Normal Q-Q Plot for GardenB")
qqline(garden$gardenB)

# Observation
# The dots are close to the plain line. Means the distribution follows the normal one.
```
```{r}
# Q-Q plot for garden C
qqnorm(garden$gardenC, main = "Normal Q-Q Plot for GardenC")
qqline(garden$gardenC)

# Observation
# The dots are not close to the plain line. Means the distribution does not follow the normal one.
```


Next step is to draw the Q-Q plot with ggplot.

```{r}
garden_2 %>%
  group_by(Garden) %>%
  mutate(scaled_ozone = scale(ozone)) %>%
  ggplot(aes(sample = scaled_ozone, colour = Garden)) +
  stat_qq(distribution = stats::qnorm) +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(Garden~.) +
  labs(title = "Normal Q-Q plot",
       x = "Theoretical Quantities",
       y = "Scaled sample Quantities")
```

I now plot everything together

```{r}
library(cowplot)

garden_2 %>%
  ggplot(aes(x = ozone, color = Garden, fill = Garden)) +
  geom_density(alpha = 0.5) +
  scale_x_continuous(limits = c(min(limits_x$min), max(limits_x$max))) + 
  theme_classic() +
  labs(title = "Garden Densities",
       x = "[ozone]",
       y = "density") -> p1

garden_2 %>%
  group_by(Garden) %>%
  mutate(scaled_ozone = scale(ozone)) %>%
  ggplot(aes(sample = scaled_ozone, colour = Garden)) +
  stat_qq(distribution = stats::qnorm) +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(Garden~.) +
  theme(legend.position = "none") +
  labs(title = "Normal Q-Q plot",
       x = "Theoretical Quantities",
       y = "Scaled sample Quantities") -> p2

plot_grid(p1, p2, nrow = 2)
```



# FINAL TEST

I use the Shapiro test for the three datasets, using  $\alpha=0.05$.

```{r}
tidy(shapiro.test(garden$gardenA))
tidy(shapiro.test(garden$gardenB))
tidy(shapiro.test(garden$gardenC))

# Conclusion
# Since the null hypothesis of this test is that our distribution is normal
# For gardenA and gardenB, the p-values, 0.703 are greater than 0.05. This means we can accept the hypothesis for normality for gardenA and gardenB.
# For gardenC, the p-value, 0.0082 is < 0.05. We reject the null hypothesis. gardenC's distribution is not normal.
```