---
title: "Studying Deforestation in Brazil"
author: "Michael Mbajwa"
date: "2021-11-19"
output: html_document
---

```{r}
library(palmerpenguins)
library(tidyverse)
library(ggplot2)
```


I am going to use a dataset from the [TidyTuesday](https://github.com/rfordatascience/tidytuesday/) initiative. Several dataset about the theme **deforestation** on April 2021 were released, we will focus on the csv called `brazil_loss.csv`. The dataset columns are described in the linked [README](https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-04-06/readme.md) and the csv is directly available at this [url](https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/brazil_loss.csv)


```{r}
url <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/brazil_loss.csv'
load_url = readr::read_csv(url)
brazil_loss <- load_url %>%
              select(!contains(c('entity', 'code')))

# Because of the nature of exploratory analysis I want to perform, I change the datatype of year column.
brazil_loss <- brazil_loss %>%
  mutate(year = as.character(year))

brazil_loss
```


```{r}
# I pivot the data to a long format to make it easier to work with
brazil_loss %>%
  pivot_longer(commercial_crops:small_scale_clearing,
               names_to = "deforestation_reasons",
               values_to = "area_ha") -> brazil_loss_long
brazil_loss_long
```



##### I plot the deforestation areas per year as bars

```{r}
# Write your answer here
brazil_loss_long %>%
  ggplot(aes(x=year,
             y=area_ha)) +
  geom_col(fill = "#f68060") + 
  theme_classic() + 
  labs(title = "Deforestation areas per year in Brazil",
       x = "Year",
       y = "Area per hectare",
       caption = "Data credit: TidyTuesday")

```


##### I repeat the same plot but fill each bar with the reasons for deforestation

```{r}
# Write your answer here
brazil_loss_long %>%
  ggplot(aes(x=year,
             y=area_ha, 
             fill = deforestation_reasons)) +
  geom_col() + 
  theme_classic() + 
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Deforestation areas per year in Brazil",
       x = "Year",
       y = "Area per hectare",
       caption = "Data credit: TidyTuesday",
       fill = "Reasons for\nDeforestation") +
  theme(plot.caption = element_text(face = "italic", hjust = 1.5))
```


##### I lump the reasons for deforestations, keeping only the top 5 reasons, lumping as "Other" the rest

```{r}
# Write your answer here
brazil_loss_long %>%
  ggplot(aes(x = year,
             y = area_ha,
             fill = fct_lump(deforestation_reasons, 5, w = area_ha) %>%
               fct_infreq())) +
  geom_col()+
  theme_classic()+
  labs(title = "Top 5 yearly causes of Deforestation in Brazil",
       fill = "Reasons for\nDeforestation",
       x = "Year",
       y = "Areas per heactare",
       caption = "Data credit: TidyTuesday") +
  theme(axis.text.x = element_text(angle = 90)) +
  theme(plot.caption = element_text(face = "italic", hjust = 1.5))
```


```{r}
# I optimize the previous plot by sorting the 5 main deforestation reasons
# Write your answer here
brazil_loss_long %>%
  group_by(year) %>%
  top_n(5, area_ha) %>%
  mutate(reasons2 = deforestation_reasons) -> top_5

brazil_loss_long %>%
  group_by(year) %>%
  left_join(top_5, by = c("year", "deforestation_reasons", "area_ha")) %>%
  replace_na(list(reasons2 = "Other")) %>%
  arrange(desc(area_ha), .by_group = TRUE) %>%
   ggplot(aes(x = year,
             y = area_ha,
             fill = fct_reorder(reasons2, area_ha, .desc = TRUE) %>%
               fct_rev())) +
  geom_col() +
  labs(title = "Top 5 yearly causes of Deforestation in Brazil",
       fill = "Reasons for\nDeforestation",
       x = "Year",
       y = "Areas per heactare",
       caption = "Data credit: TidyTuesday") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))+
  theme(plot.caption = element_text(face = "italic", hjust = 2.0))
```


